Efficiently Handling Guards in the Low Level Design of RPython's Tracing JIT

### 2012-08-10 talk/vmil2012/paper.tex
write about the size of resume data

All figures in this section do not take into account garbage collection. Pieces
of machine code can be globally invalidated or just become cold again. In both
cases the generated machine code and the related data is garbage collected. The
figures show the total amount of operations that are evaluated by the JIT and
the total amount of code and data that is generated from the optimized traces.

@@

\todo{compare to naive variant of resume data}

@@

@@

Mike Pall, the author of LuaJIT describes in a post to the lua-users mailing
list different technologies and techniques used in the implementation of
LuaJIT~\cite{Pall:2009}. Pall explains that guards in LuaJIT use a datastucture
called snapshots, similar to RPython's resume data, to store the information about
how to rebuild the state from a side-exit using the information in the snapshot
and the machine execution state. Pall also acknowledges that snapshot for
guards are associated with a large memory footprint. The solution used in
LuaJIT is to store sparse snapshots, avoiding the creation of snapshots for
every guard to reduce memory pressure. Snapshots are only created for guards
after updates to the global state, after control flow points from the original
program and for guards that are likely to fail. As an outlook Pall mentions the
plans to switch to compressed snapshots to further reduce redundancy.

>>>>

All figures in this section do not take garbage collection into account. Pieces
of machine code can be globally invalidated or just become cold again. In both
cases the generated machine code and the related data is garbage collected. The
figures show the total amount of operations that are evaluated by the JIT and
the total amount of code and data that is generated from the optimized traces.

@@

@@

Why the efficient storing of the \texttt{resume data} is a central concern in the design
of guards is illustrated by Figure~\ref{fig:backend_data}, this Figure shows
the size of the compressed \texttt{resume data}, the approximated size of
storing the \texttt{resume data} without compression and the size of
compressing the data to calculate the size of the resume data using the
\texttt{xz} compression tool, which is a ``general-purpose data compression
software with high compression ratio'' used to approximate the best possible
compression for the \texttt{resume data}.\footnote{\url{http://tukaani.org/xz/}}.

The results show that the current approach of compression and data sharing only
requires 18.3\% to 31.1\% of the space compared to the naive approach. This
shows that large parts of the resume data are redundant and can be stored more
efficiently through using the techniques described above. On the other hand
comparing the results to the xz compression which only requires between 17.1\%
and 21.1\% of the space required by our compression shows that the compression
is not optimal but a trade-off between the required space and the time needed
to build a good compressed representation of the compressed resume data for the
large amount of guards present in the traces.

@@

Mike Pall, the author of LuaJIT describes in a post to the lua-users mailing
list different technologies and techniques used in the implementation of
LuaJIT~\cite{Pall:2009}. Pall explains that guards in LuaJIT use a datastucture
called snapshots, similar to RPython's resume data, to store the information
about how to rebuild the state from a side-exit using the information in the
snapshot and the machine execution state. According to Pall~\cite{Pall:2009}
snapshots for guards in LuaJIT are associated with a large memory footprint.
The solution used in there is to store sparse snapshots, avoiding the creation
of snapshots for every guard to reduce memory pressure. Snapshots are only
created for guards after updates to the global state, after control flow points
from the original program and for guards that are likely to fail. As an outlook
Pall mentions the plans to switch to compressed snapshots to further reduce
redundancy.
